Multisensor Architecture for an Intersection Management System

### 2015-01-14 ch2-mdf_ims.tex
Updated bibliography. Changed bibliography style

\chapter [Sensor Fusion and IMS]{Multisensor Data Fusion and Intersection Management Systems}

An intersection is a highly-dynamic scenario that can be monitored using a wide range of sensors. For this reason, an efficient and accurate fusion of the information is needed. This chapter is divided into two sections. In the first section a brief overview of multisensor data fusion is presented, remarking in different architectures proposed in the literature and different approaches used for laser and video data fusion.
The second section contains a short review on intelligent transportation systems and intersection management systems, including current research on this topic. Finally, there will be presented some projects that include fusion of laser and video data for intersection managing applications.

\section {Multisensor Data Fusion}

Data fusion, also referred as mutisensor data fusion, information fusion or sensor fusion, has received several definitions from different authors in the literature. For example, Joint Directors of Laboratories defined data fusion as "multi-level, multifaceted process handling the automatic detection, association, correlation, estimation, and combination of data and information from several sources"\cite{White1991}. Luo refers to multisensor fusion and integration as "synergistic combination of sensory data from multiple sensors to provide more reliable and accurate information"\cite{Luo2002} and "to achieve inferences that are not feasible from each individual sensor operating separately"\cite{Luo2011}. Elmenreich states that sensor fusion is "the combining of sensory data or data derived from sensory data such that the resulting information is in some sense better than would be possible when these sources were used individually"\cite{Elmenreich2007}. In \cite{Bostrom2007} there is a compilation of more definitions of information fusion and the author summarize in his own statement, i.e., "Information fusion is the study of efficient methods for automatically or semi-automatically transforming information from different sources and different points in time into a representation that provides effective support for human or automated decision making".

@@

It is clear, now, that multisensor data fusion also is a multidisciplinary field, because information in a typical process, flows from sensors to applications, passing by stages of filtering, data enhancement and data extraction. Is for this that knowledge in a wide range of fields are required, e.g. signal processing, machine learning, probability and statistics, etc. Also, it would be pointless to try to define a general method, technique or architecture that fits the requirements of any system, for applying data fusion in it.

%\subsection{Overview}
\subsection{Architectures}

Although there is not a general rule of how to design or implement a sensor fusion system, many authors have proposed some architectures and guidelines for this task. One of the first proposals of fusion architecture, and probably one of the most widely used, is the JDL fusion model, originated from the US Joint Directors of Laboratories and described by Hall and Llinas in \cite{Hall1997} and \cite{Llinas1998}. The JDL fusion model was conceived to aid the developments of military applications and comprises 5 levels of data processing at which data fusion could be done. These levels and a database are connected by a bus (\ref{JDLmodel}), and are not meant to be execute sequentially and can also be executed concurrently.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.4]{fig/2/jdlmodel}
\caption{JDL fusion model. (from \cite{Hall1997}).}
\label{JDLmodel}
\end{figure}

\subsection{Laser and Video Data Fusion}

\section{Intersection Management Systems}
\subsection{Taxonomy}
\subsection{Intersection Management Systems and Multisensor Data Fusion}

@@

>>>>

%\chapter [Sensor Fusion and IMS]{Multisensor Data Fusion and Intersection Management Systems}

An intersection is a highly-dynamic scenario that can be monitored using a wide range of sensors. For this reason, an efficient and accurate fusion of the information is needed. This chapter is divided into two sections. In the first section a brief overview of multisensor data fusion is presented, remarking in different architectures proposed in the literature and different approaches used for laser and video data fusion are described.
The second section contains a short review on intelligent transportation systems and intersection management systems, including current research on this topic. Finally, some projects that include fusion of laser and video data for intersection managing applications are presented.

\section{Multisensor Data Fusion}

Data fusion, also referred as mutisensor data fusion, information fusion or sensor fusion, has received several definitions from different authors in the literature. For example, Joint Directors of Laboratories defined data fusion as "multi-level, multifaceted process handling the automatic detection, association, correlation, estimation, and combination of data and information from several sources"\cite{White1991}. Luo refers to multisensor fusion and integration as "synergistic combination of sensory data from multiple sensors to provide more reliable and accurate information"\cite{Luo2002} and "to achieve inferences that are not feasible from each individual sensor operating separately"\cite{Luo2011}. Elmenreich states that sensor fusion is "the combining of sensory data or data derived from sensory data such that the resulting information is in some sense better than would be possible when these sources were used individually"\cite{Elmenreich2007}. In \cite{Bostrom2007} there is a compilation of more definitions of information fusion and the author summarize in his own statement as follows: "Information fusion is the study of efficient methods for automatically or semi-automatically transforming information from different sources and different points in time into a representation that provides effective support for human or automated decision making".

@@

It is clear, now, that multisensor data fusion is a multidisciplinary field, because information in a typical process, flows from sensors to applications, passing through stages of filtering, data enhancement and data extraction. It is for this that knowledge in a wide range of fields are required, e.g. signal processing, machine learning, probability and statistics, etc. Also, it would be pointless to try to define a general method, technique or architecture that fits the requirements of any system, for applying data fusion in it.

%\subsection{Overview}
\subsection{Data Fusion Architectures and Models}

Although there is not a general rule of how to design or implement a sensor fusion system, many authors have proposed some models, architectures and guidelines for this task. Three well-known models are Waterfall model, JDL fusion model and Multisensor integration fusion model.

\subsubsection{Waterfall model}

Harris and Markin in \cite{Harris1998} and CITE, proposed a model named Waterfall, in which they describe the fusion process as an information flow through sensing to decision-making. They describe 3 levels of processing with 2 inner stages each ((\ref{WaterfallModel}). The first level is about transform the raw data from sensor to a better representation of the measured phenomena through signal processing and having in mind sensor models and nature of the process itself. The second level objective is to find a meaningful description of the data, reducing its volume while maximising information. This is done using feature extraction and pattern recognition techniques. The third level is the high level of the process in which situation assessment and decision making are performed, based on data available, configuration parameters, database information or human interaction. Finally, a feedback from high-level to low-level (sensor) is done, advising the whole system for re-calibration or reconfiguration.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.4]{fig/2/WaterfallModel.png}
\caption{Waterfall model. (from \cite{Esteban2005}).}
\label{WaterfallModel}
\end{figure}

\subsubsection{JDL fusion model}

One of the first proposals of fusion architecture, and probably one of the most widely used, is the JDL fusion model, originated from the US Joint Directors of Laboratories and described by Hall and Llinas in \cite{Hall1997} and \cite{Llinas1998}. The JDL fusion model was conceived to aid the developments of military applications and comprises 5 levels of data processing at which data fusion could be done. These levels and a database are connected by a bus (\ref{JDLmodel}), and are not meant to be execute sequentially and can also be executed concurrently.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.4]{fig/2/jdlmodel.png}
\caption{JDL fusion model. (from \cite{Hall1997}).}
\label{JDLmodel}
\end{figure}

The first stage, referred as level-0 is the source preprocessing in which raw data is handled to concentrate the more pertinent data for the current situation. The level-1 is for object refinement, starting with the alignment of the data in a commonly space-time reference frame. Then, performs identification and tracking of objects using different techniques. Situation refinement is at level-2, which takes observed and partially-observed object from level-1 and tries to find a contextual description between them. Level-3, threat refinement, is the level in which results from level-2 are interpreted looking for possible advantages and disadvantages for the system to operate, based on previous knowledge and predictions about executing an action.

\subsubsection{Multisensor Fusion Integration model}

Luo and Kai, in \cite{Luo1989, Luo1990}, proposed a full integration model for data fusion in which they define a three-level hierarchy for sensor fusion: data-fusion, feature-fusion and decision-fusion. This model separate MFI in five classes, based on Input/Output pair: Data in-data out fusion, data in-feature out fusion, feature in-feature out fusion, feature in-decision out fusion, and decision in-decision out fusion \cite{Luo2011} (figure \ref{fusionClasses}).

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{fig/2/fusionClasses.png}
\caption{Five classes of Multisensor Fusion.}
\label{fusionClasses}
\end{figure}

Also, they made a clear distinction between multisensor fusion and multisensor integration, being the former the process in which information provided by a set of sensors is combined in any of the three levels aforementioned, and the latter is how sensor fusion could be integrated in a full system in order to assist in a particular task. As is depicted in figure \ref{MFIArch}, sensor fusion is an element of the whole MFI architecture, which also includes block for sensor managing tasks, like control, selection and registration of sensors, previously to the fusion process. A sensor processing stage and a system controlling module are also included after the sensor fusion stage.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.4]{fig/2/MFIArch.png}
\caption{Multisensor Fusion Integration architecture (from \cite{Luo2011}).}
\label{MFIArch}
\end{figure}

\subsection{Classification of Data Fusion Architectures}

Elmenreich in \cite{Elmenreich2007} classify fusion models in three categories: Abstract models, generic architectures and rigid architectures. Abstract model are not intended to show how to implement a sensor fusion system, but to explain which processes are done in it. Generic architectures gives an outline on how a system could be implemented in an application, but do not specify what type of hardware, database or communication system could be used. Finally, rigid architectures, are a good guide for implementation  of data fusion in certain applications, at the cost that several design decisions have been already taken, making expensive the migration to another architecture.

In addition to models previously mentioned, there exists more proposal for data fusion architectures in literature, as can be viewed in figure \ref{fusionModelsClas}.

% TODO: Change for a table
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.6]{fig/2/fusionModelsClas.png}
\caption{Classification of some data fusion models}
\label{fusionModelsClas}
\end{figure}

\subsection{Algorithms in Data Fusion}

Different types of algorithms have been used in implementing data fusion systems, depending on a variety of conditions like, level of fusion, type of the data, nature of environment, etc. Constrains like processing and memory limitations, centralised or distributed schemes, human-interactive or completely autonomous process, also determine which algorithms should be used in fusion process.

Khaleghi in \cite{Khaleghi2013} describes a classification for fusion algorithms based on challenging problems that arise from the data to be fused, due to the variety of sensors and the nature of the application environment. This classification is not on the algorithms directly, but on the theory or framework in which they originate. Four types of data are enumerated: Imperfect data, correlated data, inconsistent data and disparate data. For thi

%\begin{figure}[ht!]
%\centering
%\includegraphics[scale=0.275]{fig/2/fusionAlg.png}
%\caption{Fusion algorithms based on level of fusion. (from \cite{Luo2011})}
%\label{fusionAlgs}
%\end{figure}

%\subsection{Laser and Video Data Fusion}

\section{Intersection Management Systems}

Intelligent Transportation Systems includes a wide range of applications and services transversal to many knowledge areas. For classifying those services, some taxonomies have been proposed like the ones presented in \cite[Ch.1]{Sussman2005} and \cite{Williams2008}. From described categories and classes, Advanced Traffic Management Systems have to be considered when an intelligent handling of traffic needs to be deployed.

One of the most desirable scenarios to improve efficiency and safety is an intersection. This because intersections are places where vehicles arrive from different directions at different velocities, increasing the chances for incidents and crashes. Choi \cite{Choi2010} states that 40\% of reported traffic accidents in the US, were intersection related. Also, in  \cite{CorporacionFondodePrevencionVial2010}, is reported that for Colombia in 2011, most of the accidents in the main cities were at intersections.

%El concepto de sistemas inteligentes de transporte (ITS), presenta un amplio campo de acci\'{o}n transversal a diferentes \'{a}reas del conocimiento e igualmente presenta una gran variedad de aplicaciones y servicios. Para la clasificación de estos servicios se han definido diferentes taxonom\'{i}as como las descritas en \cite[Ch.1]{Sussman2005} y en \cite{Williams2008}. De las clases y categor\'{i}as descritas, se tiene un subgrupo de los ITS que debe ser contemplado para cumplir con el objetivo de este proyecto, que es el que incluye los servicios y sistemas para la administraci\'{o}n y operaci\'{o}n del tr\'{a}fico (o Sistemas Avanzados de Administraci\'{o}n de Tr\'{a}fico, del ingl\'{e}s ATMS, Advanced Traffic Management Systems).
%
%Uno de los escenarios donde se busca mejorar la eficiencia y la seguridad usando estos sistemas son las intersecciones viales, ya que son puntos donde se encuentran veh\'{i}culos en diferentes direcciones a diferentes velocidades, lo cual incrementa la probabilidad de incidentes y choques. Cerca del 40\% de los accidentes de tr\'{a}nsito reportados en Estados Unidos en 2008, estaban relacionados con intersecciones \cite{Choi2010}. En Colombia, para el 2011 la mayor\'{i}a de accidentes en las principales ciudades del pa\'{i}s se concentraba en intersecciones \cite{CorporacionFondodePrevencionVial2010}.

%\section[Types of IMS]{Types of Intersection Management Systems}

Different types of applications and systems appear in order to address this issues. Some tasks these systems realize are intersection monitoring, vehicles detection, incident warning, collision avoidance, among others. Communication approaches in these systems are based in the interaction between vehicles and infrastructure, commonly divided in three groups: Vehicle-to-vehicle communication (V2V) , Vehicle-to-infrastructure communication (V2I) and the combination of both of them (V2X).

%Para afrontar esta problem\'{a}tica aparecen diferentes tipos de aplicaciones y sistemas para realizar tareas como monitoreo y administraci\'{o}n de intersecciones, detecci\'{o}n de veh\'{i}culos, advertencia de incidentes, prevenci\'{o}n de colisiones, entre otros. El esquema de comuniciaci\'{o}n de estos sistemas se basa en la interacci\'{o}n entre los veh\'{i}culos y la infraestructura, com\'{u}nmente dividida en 3 tipos: Comunicaci\'{o}n veh\'{i}culo a veh\'{i}culo (V2V), veh\'{i}culo a infraestructura (V2I) y la combinaci\'{o}n de las dos anteriores (V2X).

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.75]{fig/2/fig1}
\caption{Generic block diagram of an Intersection Management System (from \cite{Chan2005}).}
\label{arch}
\end{figure}

Figure \ref{arch}, presents a generic block diagram of an Intersection Management System composed by three main components: (a) Data source, that could be infrastructure sensors, like inductive loops, range sensors or cameras, and also could be vehicle sensors. (b) Decision system, which is the core of the whole system, is in charge of analyse and process information provided by infrastructure, vehicles and environment in order to predict future incidents, control traffic and generate safe decisions and warnings alerts. Finally, (c) is the presentation and displaying of the output of decision system. It could be at the infrastructure using dynamic signals, semaphore controlling, or vehicle-directed through on-board visualization/notification system.

%En la figura \ref{arch}, se presenta un diagrama de bloques general de un sistema de administraci\'{o}n de una intersecci\'{o}n en donde se destacan 3 componentes: (a) Las fuentes de los datos, que pueden ser sensores en la infraestructura, como sensores inductivos, sensores de rango, c\'{a}maras, o sensores en el veh\'{i}culo. (b) Sistema  de decisi\'{o}n o administraci\'{o}n, que es el encargado de analizar la informaci\'{o}n de los veh\'{i}culos y el entorno, predecir incidentes futuros y enviar se\~{n}ales de alerta. y (c) La presentaci\'{o}n de informaci\'{o}n de control, que puede ser en la infraestructura con se\~{n}ales din\'{a}micas, control de los sem\'{a}foros, como tambi\'{e}n puede ser en el veh\'{i}culo mediante un dispositivo de visualizaci\'{o}n.

\subsection{Applications in IMS}

Intersection monitoring is a required task to be done within intelligent transportation systems for high-level applications as traffic analysis, counting and classification of vehicles or pedestrians, events and incident detection, accidents prevention and security and surveillance applications.

To complete these tasks, many approaches have been proposed for information capture stage, both taking data from vehicle or from the infrastructure. For example, monitoring mobile wireless network like Bluetooth \cite{Friesen2013}, detecting vehicle presence using inductive loops or RFID tags, identifying vehicles and pedestrian using cameras \cite{Buch2011, Strigel2013}, single or multiple laser-scanners \cite{Meissner2010, Meissner2012, Zhao2006, Zhao2008, Zhao2012}, and multi-sensor integrated systems \cite{Meissner2013, Meissner2013a, Meissner2014, GoldHammer2012, Zhao2009}.

Other application are traffic-control oriented, rather than to sensing and data-processing stage. The aim of these applications are to maximise vehicle flow, reduce accidents and improve safety. Many works have been done and although its objective is the same, are diversity between them regarding its communication approach (V2V, V2I or V2X), sensors and information that are used, the execution scenario (simulation, real implementation or augmented reality) and the nature of vehicle driving (Human driver or autonomous vehicles).

%El monitoreo en las intersecciones viales es una tarea requerida en los sistemas inteligentes de transporte para aplicaciones de alto nivel como  an\'{a}lisis de tr\'{a}fico, conteo y  clasificaci\'{o}n de veh\'{i}culos, detecci\'{o}n de eventos, prevenci\'{o}n de accidentes y aplicaciones de seguridad y vigilancia. Para cumplir con estas tareas se han propuesto diversas soluciones para obtener la informaci\'{o}n del entorno, ya sea proveniente de los veh\'{i}culos, por ejemplo monitoreando conexiones a redes m\'{o}viles, como Bluetooth \cite{Friesen2013}, detectando su presencia usando sensores inductivos o etiquetas RFID, y tambi\'{e}n tomando datos desde la infraestructura, usando c\'{a}maras  \cite{Buch2011}, sensores l\'{a}ser o arreglos de estos \cite{Zhao2012} o sistemas que integran m\'{u}ltiples tipos de sensores \cite{Zhao2009}\cite{Pyykonen2010}. Otras aplicaciones adem\'{a}s de monitorear y supervisar las intersecciones, est\'{a}n orientadas al control del tr\'{a}fico en estas con  el objetivo de maximizar el flujo, disminuir la accidentalidad y aumentar la seguridad. Diversos trabajos se han hecho y aunque su finalidad es similar, existe variedad entre s\'{i}, ya sea seg\'{u}n el tipo de comunicaci\'{o}n en la que se basan (V2V, V2I o V2X), el tipo de sensores e informaci\'{o}n que usan, el entorno de ejecuci\'{o}n (simulaci\'{o}n, real o realidad aumentada) y el tipo de veh\'{i}culos (con conductores humanos o veh\'{i}culos aut\'{o}nomos).

In \cite{Ball2010} and \cite{Azimi2012}, intervehicular-communication based architectures are proposed. In these, each vehicle send to the others its own information about position, velocity, destination and more, and then they coordinate how the access to the intersection should be done. In \cite{Guerrero-Ibanez2013}, is presented a system capable of work in V2I mode, additional to V2V, using inductive loops and RFID information for detecting vehicles. The information is gathered by an intersection controller which manage the access to the crossroad. Another V2I based systems is presented in \cite{Basma2011}, where wireless magnetic nodes are used to detect presence and to send data to a base station, which determines possible collisions and warns driver through a visualisation system in infrastructure. Furthermore, in \cite{Dresner2008} and \cite{CondeBento2012}, are presented simulation-based systems for traffic controlling, using agents and applied to autonomous vehicles. In \cite{Quinlan2010}, an augmented reality implementation of the system proposed in \cite{Dresner2008} is done. In this work, approaching vehicles request to an intersection manager for authorisation to cross and is the manager which decides if it is safe to cross or not, depending on the defined traffic policies.

%En \cite{Ball2010} y \cite{Azimi2012} se proponen arquitecturas basadas en comunicaci\'{o}n intervehicular, donde cada auto env\'{i}a a los dem\'{a}s informaci\'{o}n correspondiente a su posici\'{o}n, velocidad, destino y otros datos, y as\'{i} coordinan el acceso a la intersecci\'{o}n. En \cite{Guerrero-Ibanez2013} se presenta un sistema que adem\'{a}s de trabajar en modo V2V, presenta la opción V2I que incluye sensores inductivos y m\'{o}dulos RFID para detectar veh\'{i}culos. La informaci\'{o}n adquirida es procesada por un controlador de intersecci\'{o}n, el cu\'{a}l se encarga de administrar el ingreso al cruce vial. Otro sistema basado en V2I es \cite{Basma2011} donde tambien se usan nodos de sensores magnéticos inalámbricos que se comunican con una estación base, la cual mediante algoritmos de predicción determina posibles colisiones y advierte a los conductores mediante un sistema de visualización en la infraestructura. Por otro lado, \cite{Dresner2008} y \cite{CondeBento2012} presentan aplicaciones en simulación para el control de las intersecciones usando agentes y aplicado en vehículos autónomos. En \cite{Quinlan2010} se hace una implementación en realidad aumentada del sistema propuesto por Dresner y Stone \cite{Dresner2008}. En esta implementación, los vehículos que se aproximan a la intersección solicitan autorización al controlador para ingresar y es este quien decide cuando pueden cruzar, dependiendo de las politicas de administración definidas.

%\subsection{Taxonomy}

%\subsection{Intersection Management Systems and Multisensor Data Fusion}

@@

Several models, architectures and frameworks have been proposed in literature; some of them to show how a data fusion system should work and some other providing guidelines on how to implement it in a given application.
